\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-03-20 doc/sw-system.tex
Up to section 2.3.2 (stage 2) in AV comments #96

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Space Warps System Paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

A screenshot of the \sw classification interface (CI) is shown in
\Fref{fig:screenshot}. The CI is the centrepiece of the \sw website,
\texttt{http://spacewarps.org}; the web application is written in
coffeescript, css and html and follows the general design of others written by
the Zooniverse team.\footnote{The \sw web application code is open source and
can be accessed from \texttt{https://github.com/zooniverse/Lens-Zoo}}  The
focus of the CI is a large display of the current pre-prepared PNG image of
the ``subject'' being inspected.   When the image is clicked on by the
volunteer, a marker symbol appears where the pointer was. Several markers can
be placed.  The next image moves rapidly in from a queue formed at the right
hand side of the screen when the ``Finished marking'' button is pressed. At
the same time, the positions of the markers are written out to the
classification database, in an entry that also stores the ID of the subject,
the username (or IP address) of the volunteer, a timestamp and some other
metadata.

Gravitational lenses are rare: typically, most of the images will not contain
a lens candidate, and these need to be quickly rejected by the inspector. The
queue allows several images to be pre-loaded while the volunteer is
classifying the current subject, and the rapid movement is designed to
encourage volunteers to classify rapidly.

% PJM: how quickly? How much clock time per classification?

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. The first is the ``Quick Dashboard'' (QD) a
more advanced image viewer. This allows the viewer to compare three different
contrast and color balance settings, to help bring out subtle features, and to
zoom in on interesting regions of the image to assess small features. Markers
can be placed in the Quick Dashboard just the same as in the main CI image
viewer. The second is a link to that subject's page in the project discussion
forum, \texttt{http://talk.spacewarps.org} (known as ``Talk''). Here,
volunteers can discuss the features they have seen either before they submit
their classification, or after, if they ``favorite'' the subject. There is no
``back'' button: each volunteer may only classify a given subject once.
However, the presence of an option to see what others think about any given
subject before submitting your own classification means that the
classifications may not be  strictly independent; the advantage of this system
is that volunteers can learn from others what constitutes a good lens
candidate. In practice, we might expect this to be a relatively unimportant
educational resource, given the explicit training we provide for the
volunteers, which we describe in the next section.

@@

Gravitational lenses are unfamiliar objects to volunteers who are new
to the site. New volunteers need to learn what lenses look like as quickly as
possible, so that they can contribute informative classifications. They also
need to learn what lenses do not look like, in order to reduce the false
positive detection rate. There are three primary mechanisms in the \sw system
for teaching the volunteers what to look for. These are, in the order in which
they are encountered, an inline tutorial, instant feedback,
and a ``Spotter's Guide.''

@@

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. Training continues beyond the first image tutorial through
``training subjects'' inserted randomly into the stream.  These training
subjects are either simulated lenses (known as ``sims''), or survey images
that were expert-classified (by AV, AM and PM) and found not to contain any
lens candidates (these images are known as ``duds''). The tutorial explains
that the volunteers  will be shown such training images. They are also
informed that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either positive feedback for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or negative feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'')

The initial frequency of the training images is set to be two in five;
subjects are drawn randomly from the pool of training images with this
frequency. The pool contains equal numbers of sims and duds, and the draw is
made without replacement (for that volunteer). As the number of
classifications made by a volunteer increases, this frequency is decreased, to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 0.3$ for the second 20
subjects, $0.2$ for the third 20 subjects, and so on).

@@

In \Fref{fig:training-gallery} we show 8 example training images from the
first \sw project (\Sref{sec:data} below).

@@

We now describe briefly the two-stage strategy that was employed in the CFHTLS
project, initial classification (involving the rejection of very large numbers
of non-lenses), and refinement (to further narrow down the sample). The web
application was reconfigured between the two stages, to assist in their
functioning.

@@

The goal of a stage 1 classification is to achieve a high rejection rate,
while maintaining high completeness.  In this mode, therefore, the pre-loading
of images was used to make the sliding in of new subjects happen quickly, to
provide a sense of urgency: initial classification must be done fairly
quickly for the search to be completed within a reasonable time period.  We
expect some trade-off between speed and accuracy, which we return to in the
results section below. Completion of the search requires subjects to be
``retired'' over time, as a result of their being classified. We do this by
analyzing the classifications on a daily basis, as described in
\Sref{sec:swap} below. As subjects are retired, new ones are ingested into
the web app for classification. This means that the discovery of lens
candidates in stage 1 is truly a community effort: to detect a lens candidate,
many non-lenses must first be rejected.

The stage 1 training set was chosen to be quite clear cut, in order to err on
the side of inclusivity, and so ensure high completeness. For the training
duds we selected several hundred images at random, and three of us (AV, AM,
PM) inspected them and discarded anything that could be a lens candidate. The
remainder were ingested into the site.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Space Warps System Paper
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

A screenshot of the \sw classification interface (CI) is shown in
\Fref{fig:screenshot}. The CI is the centrepiece of the \sw website,
\texttt{http://spacewarps.org}; the web application is written in coffeescript,
css and html and follows the general design of others written by the Zooniverse
team.\footnote{The \sw web application code is open source and can be accessed
from \texttt{https://github.com/zooniverse/Lens-Zoo}}  The focus of the CI is a
large display of a pre-prepared PNG image of the ``subject'' being inspected.
When the image is clicked on by the volunteer, a marker symbol appears where the
pointer was. Several markers can be placed.

The next image moves rapidly in from a queue formed at the right hand side of
the screen when the ``Finished marking'' button is pressed. Gravitational lenses
are rare: typically, most of the images will not contain a lens candidate, and
these need to be quickly rejected by the inspector. The queue allows several
images to be pre-loaded while the volunteer is classifying the current subject,
and the rapid movement is designed to encourage volunteers to classify rapidly.
After each classification, the positions of the markers are written out to the
classification database, in an entry that also stores the ID of the subject, the
username (or IP address) of the volunteer, a timestamp and some other metadata.

For the more interesting subjects, the CI offers two features that  enable
further investigation of the subjects. The first is the ``Quick Dashboard'' (QD)
a more advanced image viewer. This allows the viewer to compare three different
contrast and color balance settings, to help bring out subtle features, and to
pan and zoom in on interesting regions of the image to assess small features.
Markers can be placed in the Quick Dashboard just the same as in the main CI
image viewer. The second is a link to that subject's page in the project
discussion forum, \texttt{http://talk.spacewarps.org} (known as ``Talk''). Here,
volunteers can discuss the features they have seen either before they submit
their classification, or after, if they ``favorite'' the subject. There is no
``back'' button: each volunteer may only classify a given subject once. However,
the presence of an option to see what others think about any given subject
before submitting your own classification means that the classifications may not
be strictly independent; the advantage of this system is that volunteers can
learn from others what constitutes a good lens candidate. This is not, however,
the primary educational resource; we describe the explicit training that we
provide for the volunteers in the next section.

@@

Gravitational lenses are unfamiliar objects to volunteers who are new
to the site. New volunteers need to learn what lenses look like as quickly as
possible, so that they can contribute informative classifications. They also
need to learn what lenses do not look like, in order to reduce the false
positive detection rate. There are three primary mechanisms in the \sw system
for teaching the volunteers what to look for. These are, in the order in which
they are encountered, an inline tutorial, instant feedback,
and a ``Spotter's Guide.'' As well as this, we provide ``About'', ``FAQ'' and ``Science'' pages explaining the physics of gravitational lensing. While we expect that the insight from this static material should help volunteers make sense of the features in the images, we focus on the more dynamic, activity-based training early, when engaging new volunteers to participate takes priority.

@@

The second image viewed after the initial tutorial image is already a survey
image, in order to get the volunteers engaged in the real task as quickly as
possible. Training continues beyond the first image tutorial through
``training subjects'' inserted randomly into the stream.  These training
subjects are either simulated lenses (known as ``sims''), or survey images
that were expert-classified and found not to contain any
lens candidates (these images are known as ``duds''). The tutorial explains
that the volunteers  will be shown such training images. They are also
informed that they will receive instant feedback about their performance after
classifying (blind) any of these training subjects. Indeed, after a volunteer
finishes marking a training subject and hits ``Finished marking,'' a pop-up
message is generated, containing either positive feedback for a successful
classification (for example, ``Well done! You spotted a simulated lens,'' as
in \Fref{fig:screenshot}) or negative feedback for an unsuccessful one (for
example, ``There is no gravitational lens in this field!'').

\todo{PJM}{Insert simulations paragraph here. General description of sim generation. Principles: use real images, draw from realistic populations, strike balance between easy to see (for education) and realistically difficult (to push sensitivity). \#106}.

When a volunteer begins classifying, they are initially shown training images at
a frequency of two in five; these subjects are drawn at random from a pool
containing equal numbers of sims and duds, such that that volunteer never sees a given image more than once. As the number of classifications made by a
volunteer increases, this frequency is decreased, to
$2/(5\times2^{(\textrm{int}(N_c/20)+1)/2})$ ($\approx 0.3$ for the second 20
subjects, $0.2$ for the third 20 subjects, and so on).

@@

In Figures~\ref{fig:training-gallery:sims}~and~\ref{fig:training-gallery:duds}
we show example training images from the first \sw project (\Sref{sec:data}
below).

@@

We now describe briefly the two-stage strategy that was employed in the CFHTLS
project: initial classification (involving the rejection of very large numbers
of non-lenses), and refinement (to further narrow down the sample). The web
application was reconfigured between the two stages, to assist in their
functioning.

@@

The goal of a stage 1 classification is to achieve a high rejection rate,
while maintaining high completeness.  In this mode, therefore, the pre-loading
of images was used to make the sliding in of new subjects happen quickly, to
provide a sense of urgency: initial classification must be done fairly
quickly for the search to be completed within a reasonable time period.  We
expect some trade-off between speed and accuracy, which we return to in the
results section below. Completion of the search requires subjects to be
``retired'' over time, as a result of their being classified. We did this by
analyzing the classifications on a daily basis, as described in
\Sref{sec:swap} below. As subjects were retired, new ones were ingested into
the web app for classification. This means that the discovery of lens
candidates in stage 1 is truly a community effort: to detect a lens candidate,
many non-lenses must first be rejected, and several classifications by different inspectors are needed in either case.

The stage 1 training set was chosen to be quite clear cut, in order to err on
the side of inclusivity, and so ensure high completeness. For the training
duds we selected several hundred images at random, and three of us (AM, with AM and PM) inspected them and discarded anything that could be considered a lens candidate.

% PJM IS HERE
